---
title: "Demo distributed model"
author: "Willem Vervoort & Rafael Navas"
date: "9 July 2018"
output: pdf_document
---
```{r setup, echo=F, warning=F, message=F}
# root dir
knitr::opts_knit$set(root.dir = "C:/Users/rver4657/owncloud/Uruguay/coursematerial")
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(lubridate)
require(hydromad)
require(knitr)
```

Set the working directory
```{r, eval=F}
 setwd("C:/Users/rver4657/Dropbox (Sydney Uni)/IMFIA_course_July2018")
```

Load packages

```{r}
require(tidyverse)
require(lubridate)
require(hydromad)
```
# Defining the model

## Loading data and model functions

Source the modelling functions

```{r}
source("scripts/GR4J@distr.R")
```

Load the data
```{r}
load("data/SL_distdata.RDATA")
```

Show the data
```{r}
head(SL_distdata)
```

show the parameters of the GR4J

```{r}
# for subbasin 1:
All_par$Sub1.par
names(All_par)
```
Parameters of the reaches

```{r}
# for subbasin 1:
All_reach_par$reach1.par
names(All_reach_par)
```

## Building the model

Define the data window
```{r}
# Use data for 2000 - 2005
Data_in <- window(SL_distdata,
            start = "2000-01-01",
            end ="2005-12-31")
```

Define and run the model

```{r}
SLrun = GR4JSubBasins.run(sb=3,
                          order = c(1,2,3),
                          Data = Data_in,
                          spar = All_par,
                          rpar = All_reach_par,
                          sbnames = c("PasoTroncos",
                          "FrayMarcos",
                          "PasoPache"),
                          return_state = T)
```

plot the result of the function

```{r}
xyplot(SLrun)
```

# Optimisation

Five steps:  
- Create vector x  
- Create vector par
- Create vector calibrate_on
- Data and model_input  
- Create function for objective function

## step 1 create x

```{r}
# initial guesses, using the SubX.par in All_par. and
# reachX.par in All_reach_par, where X is a number
x <- c(All_par$Sub1.par[2:5], All_par$Sub2.par[2:5],
All_par$Sub3.par[2:5], All_reach_par$reach1.par,
All_reach_par$reach2.par)
str(x)
```

## step 2 define parameters

```{r}
# all the parameters of the model
par_in <- c(All_par$Sub1.par, All_par$Sub2.par,
All_par$Sub3.par, All_reach_par$reach1.par,
All_reach_par$reach2.par)
str(par_in)
```

## step 3 vector to match positions from par to x

```{r}
# define which parameters to calibrate
# We want to
Fit_these <- c(2:5,7:10,12:19)

str(par_in[Fit_these])
```

## step 4 Data and the model input  
We already have the data: Data_in.

```{r}
# Use data for 2000 - 2005
# created earlier as Data_in
# define the model_input (optional, as these are the defaults)
model_input_in <- list(sb = 3, order = c(1,2,3),
                          sbnames = c("PasoTroncos",
                          "FrayMarcos",
                          "PasoPache"))
```

## step 5 define the objective function

added this (you don't need this explicitly)

```{r}
DistGR4J_objfun(x = x, 
      calibrate_on = Fit_these, # which positions of parameters
      parM = par_in, # all parameters
      Data = Data_in, # The input data (Q,P,E)
      model_input = model_input_in, # (opt) rest model input
      objective = "r.squared", # objective function(optional)
      weights = c(1,1,1)) # weights (optional)

```

## run optim

```{r}
Fit_SL <- optim(par = x, # values of the parameters to fit
        DistGR4J_objfun, # function to fit
        calibrate_on = Fit_these, # which positions of parameters
        parM = par_in, # all parameters
        Data = Data_in, # The input data (Q,P,E)
        model_input = model_input_in, # (opt) rest model input
        objective = "r.squared", # objective function(optional)
        weights = c(1,1.5,2), # weights (optional)
        method="Nelder-Mead") # options for optim

Fit_SL$par
```

## Analyse the results

Write the new parameters into the model parameter vectors

Use the rewrite_pars()

```{r}
# insert fitted values into parameters
# using rewrite_pars()
# this creates a list of reach and sub parameters
new_pars <- rewrite_pars(All_par,All_reach_par,Fit_SL$par)
str(new_pars)
```

Now you can rerun the model with the new parameters

```{r}
Fitted_SL <- GR4JSubBasins.run(sb=3,
                          order = c(1,2,3),
                          Data = Data_in,
                          spar = new_pars$sub,
                          rpar = new_pars$reach,
                          sbnames = c("PasoTroncos",
                          "FrayMarcos",
                          "PasoPache"))

xyplot(Fitted_SL)
```

But we want to compare to the observed values. Use utility function plot_results()

```{r}
# plot using `plot_results()`
results <- plot_results(Fitted_SL,Data_in)
```

What are the stats of the fit:
use stats_fun() another utility function

```{r}
# using stats_fun on results
stats_fun(results, decimal=2)
```








