---
title: "Part3OptimisationHydromad"
author: "Willem Vervoort"
date: "3 July 2018"
output: pdf_document
---

This is the document that gives the code that was done in class for the third part of day 1

```{r setup, echo = F}
# root dir
knitr::opts_knit$set(root.dir = "C:/Users/rver4657/owncloud/Uruguay/coursematerial")
```

# Start of real code

You might want to set your working directory

```{r}
setwd("C:/Users/rver4657/owncloud/Uruguay/coursematerial")
```

packages
```{r}
require(hydromad)
```

Load the data
```{r}
data(Cotter)
# show first 6 rows
head(Cotter)
```

```{r}
# Make a plot of the data
xyplot(Cotter)
```

An example Hydromad structure

```{r}
# specify a model structure (GR4J) with some arbritrary parameters
CMod <- hydromad(Cotter[1:1000,], sma="gr4j",
                 routing="gr4jrouting",etmult=0.15,
                 x1 = 665, x2 = 10, x3 = 90,
                 x4 = 3.8)

```

You can plot what this model object would predict

```{r}
xyplot(predict(CMod, return_state = TRUE,
  return_components = TRUE),
  strip = FALSE, strip.left = TRUE)
```
Print
```{r}
print(CMod)
```
Coefficients
```{r}
coefficients(CMod)
```

# Optimisation with hydromad

Split the data using window()
```{r}
# split the data to use for calibration
Data_Cal<- window(Cotter, start = "1970-01-01",
                  end = "1975-12-31")
```

Redfine the model with ranges for the parameters

```{r}
CMod <- hydromad(Data_Cal,
  sma="gr4j", routing="gr4jrouting",
  etmult=c(0.05,0.5),x1 = c(500,3000), 
  x2 = c(-3,20),
  x3 =c(5,500), x4 = c(0.5,10))
```

Use fitByOptim() to fit the model

```{r}
CotterFit <- fitByOptim(CMod,
  objective=~hmadstat("r.squared")(Q,X),
  samples=1000,method="PORT")
```
Ask for summary

```{r}
summary(CotterFit)
```

Plot the fit

```{r}
xyplot(CotterFit, with.P=TRUE,
       xlim=as.Date(c("1974-01-01", "1975-12-31")))
```






